MSF_Groupwork-NBA_shotlog

### 2016-10-11 MSFGroupwork-NBAHotHand.Rmd
fixed mistake - hot hand theory

## Installing libraries

```{r}
install.packages('reshape')

## Calling libraries
```{r}
library(reshape)
library(ggplot2)

@@

### Home Advantage analysis

```{r}
# replacing W/L in column "W" with 1/0 (1 for win, 0 for loss)
df[,"W"] <- as.character(df[,"W"])

@@

#Close Game Effect (Justin)

```{r}
# create new variable to seperate overtime

df[, "OVERTIME"] <- as.numeric()

for (i in 1:nrow(df)) {
  if (df[i, "PERIOD"] >= 5) df[i, "OVERTIME"] <- 1
  if (df[i, "PERIOD"] <= 4) df[i, "OVERTIME"] <- 0
}

# accuracy 
acc_ot <- as.numeric
acc_ot <- sum(df$OVERTIME == 1 & df$FGM == 1)/sum(df$OVERTIME == 1)

acc_qrt1 <- as.numeric
acc_qrt1 <- sum(df$PERIOD == 1 & df$FGM == 1)/sum(df$PERIOD == 1)

acc_qrt2 <- as.numeric
acc_qrt2 <- sum(df$PERIOD == 2 & df$FGM == 1)/sum(df$PERIOD == 2)

acc_qrt3 <- as.numeric
acc_qrt3 <- sum(df$PERIOD == 3 & df$FGM == 1)/sum(df$PERIOD == 3)

acc_qrt4 <- as.numeric
acc_qrt4 <- sum(df$PERIOD == 4 & df$FGM == 1)/sum(df$PERIOD == 4)

acc_qrt5 <- as.numeric
acc_qrt5 <- sum(df$PERIOD == 5 & df$FGM == 1)/sum(df$PERIOD == 5)

acc_qrt6 <- as.numeric
acc_qrt6 <- sum(df$PERIOD == 6 & df$FGM == 1)/sum(df$PERIOD == 6)

acc_qrt7 <- as.numeric
acc_qrt7 <- sum(df$PERIOD == 7 & df$FGM == 1)/sum(df$PERIOD == 7)

acc_qrt1_4 <- as.numeric
acc_qrt1_4 <- sum(df$PERIOD <= 4 & df$FGM == 1) / sum(df$PERIOD <= 4)

t.test(df$FGM[df$PERIOD <= 4], df$FGM[df$PERIOD >= 5])
t.test(df$FGM[df$PERIOD == 1], df$FGM[df$PERIOD == 2])
t.test(df$FGM[df$PERIOD == 2], df$FGM[df$PERIOD == 3])
t.test(df$FGM[df$PERIOD == 3], df$FGM[df$PERIOD == 4])
t.test(df$FGM[df$PERIOD == 1], df$FGM[df$PERIOD == 3])
t.test(df$FGM[df$PERIOD == 2], df$FGM[df$PERIOD == 4])

# accuracy for Shot Clock

  # dealing with missing values
df[is.na(df)] <- 99

acc <- data.frame()
acc[, "SECONDS_LEFT"] <- c(1:23)

#accuracy for every second
for (i in 1:23){
  acc[i, "Accuracy"] <- (sum(df$SHOT_CLOCK > i & df$SHOT_CLOCK <= i+1 & df$SHOT_CLOCK != 99 & df$FGM == 1)/ sum(df$SHOT_CLOCK > i & df$SHOT_CLOCK <= i+1 & df$SHOT_CLOCK != 99))}

ggplot(acc, aes(x = SECONDS_LEFT, y = Accuracy)) + geom_point() + geom_smooth()

@@

shotresult

@@

player_and_game <- function(x, y) {
   player <- df[df[ ,'player_id'] == x & df[ ,'GAME_ID1'] == y, 'FGM']
   number_of_streaks(as.vector(player))
}
player_and_game(202681, 21400681)

ourtable <- head(pttsum2,10)

list_per_player <- list()
j <- 1
for(i in ourtable$player_id) {
   streak_per_player[[j]] <- player_and_game(i, ourtable[ourtable[ ,'player_id'] == i, 'GAME_ID1'])
   j <- j + 1
}
streak_per_player

>>>>

##Installing packages and calling libraries
```{r}
install.packages('reshape')
library(reshape)
library(ggplot2)

@@

```
### Home Advantage analysis
```{r}
# replacing W/L in column "W" with 1/0 (1 for win, 0 for loss)
df[,"W"] <- as.character(df[,"W"])

@@

@@

@@

#Create a function that gives us a vector of streaks for a specific player and game
player_and_game <- function(x, y) {
   player <- df[df[ ,'player_id'] == x & df[ ,'GAME_ID1'] == y, 'FGM']
   number_of_streaks(as.vector(player))
}

#Set a new data.frame for the 10 players with best performances(higher points)
ourtable <- head(pttsum2,10)

#Create a list that gives us the streaks of each player and game (10 best performances)
streak_per_player <- list()
j <- 1
for(i in ourtable$player_id) {
   streak_per_player[[j]] <- player_and_game(i, ourtable[ourtable[ ,'player_id'] == i, 'GAME_ID1'])
   j <- j + 1
}
streak_per_player

#Create a data frame for the strikes of each one of the 10 players (cannot create a data.frame including all of them, because each list has different length)
player1_df <- data.frame('streaks' = streak_per_player[[1]])
player2_df <- data.frame('streaks' = streak_per_player[[2]])
player3_df <- data.frame('streaks' = streak_per_player[[3]])
player4_df <- data.frame('streaks' = streak_per_player[[4]])
player5_df <- data.frame('streaks' = streak_per_player[[5]])
player6_df <- data.frame('streaks' = streak_per_player[[6]])
player7_df <- data.frame('streaks' = streak_per_player[[7]])
player8_df <- data.frame('streaks' = streak_per_player[[8]])
player9_df <- data.frame('streaks' = streak_per_player[[9]])
player10_df <- data.frame('streaks' = streak_per_player[[10]])
ourtable

#Visualisation for 10 players - distribution of each players' streaklengths
ggplot(data = player1_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[1,'player_name'])
ggplot(data = player2_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[2,'player_name'])
ggplot(data = player3_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[3,'player_name'])
ggplot(data = player4_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[4,'player_name'])
ggplot(data = player5_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[5,'player_name'])
ggplot(data = player6_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[6,'player_name'])
ggplot(data = player7_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[7,'player_name'])
ggplot(data = player8_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[8,'player_name'])
ggplot(data = player9_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[9,'player_name'])
ggplot(data = player10_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[10,'player_name'])

#We can see that the distribution of all players is right skewed and that all of them (especially player3) have some long shooting streaks. To prove/disprove the hot hand theory we need to examine the indipendence of the shots. If each shot that a player takes is independent of the next shot, then a missed/made shot will not affect the propability of succesfulness of the next shot. So, if consecutive shots are independent, we disprove hot hand theory, while if consecutive shots are dependent, we prove the hot hand theory.
#Simulation
shot_results <- c('made', 'missed')
simul_shooter <- sample(shot_results, size = nrow(player1_df), replace = TRUE, prob = c(0.45, 0.55))
simul_shooter

simul_shooter <- as.list(simul_shooter)
for(i in 1:length(simul_shooter)) {
   if(simul_shooter[i] == 'made') simul_shooter[i] <- 1
   else if(simul_shooter[i] == 'missed') simul_shooter[i] <- 0

}
simul_shooter
number_of_streaks(simul_shooter)
```
